Dark Control: The Default Mode Network as a Reinforcement Learning Agent

### 2018-08-10 paper/paper.tex
interm3

RL is an area of machine learning concerned with searching optimal behavioral strategies through interactions
with an \textit{environment} with the goal to maximize some \textit{cumulative reward}.
The optimal behavior typically
takes the future into account as some rewards could be \textit{delayed}.
Through repeated action on and feedback from the environment,
the agent learns how to reach goals and continuously improve the collection of reward signals
in a trial-and-error fashion (Fig. \ref{fig:rl}).
At a given moment, each taken \textit{action} $a$ triggers a change
in the \textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
$r = r(s, a,s')$ obtained by the agent.
If the collected reward outcome yields a negative value it can be
more naturally interpreted as \textit{punishment}.
In this setting, the environment is partially controlled by
the action of the agent and the reward can be thought
of as satisfaction \textemdash or aversion \textemdash that accompany the execution of
a particular action.
% Action may be delayed to achieve substantial increases in expected reward
% that is expected to grow with time.

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory input
\citep{starkweather2017dopamine}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.
In short, DMN implication in a diversity of
particularly sophisticated human behaviors
can be parsimoniously explained as instantiating probabilistic
simulations of experience coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time scales.
Such a purposeful optimization objective
may be subserved by a stochastic approximation
based on a brain implementation for Monte Carlo sampling of events and outcomes.

@@

\subsection{Markov decision processes}
RL has had considerable success in modeling many real-world problems, including super-human performance in complex video games \citep{mnih2015},
robotics \citep{ng2004,abbeel2004},
and strategic board games like the breakthrough results upon recently on the game of Go \citep{silver2016mastering} considered to be a milestone benchmark in artificial intelligence.
% We emphasize that the brain in general, and the DMN in particular,
% is a physical system governed by the laws of
% physics and can be formally described
% by Markov processes at a sufficiently coarse scale.
% It has indeed been previously proposed \citep{tegmark2016improved} that
% any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
% step is sufficiently short.
In artificial intelligence and machine learning, a popular computational model for
multi-step decision processes in such an environment are
MDPs \citep{sutton1998reinforcement}.
% For the remainder of this paper, for the sake of simplicity
% we use the abbreviation ``MDP'' to refer to the special case
% of \textit{Partially Observable Markov Decision Processes}.
An MDP operationalizes a sequential decision process
in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state.
Instead, the agent tries to optimize a \textit{subjective} reward
signal (i.e., likely to be different for another agent in the same state),
by maintaining probability distributions over actions according to their expected utility.
% This is performed alongside each transition in the state space of the environment
%  and maintains a probability distribution over the set of possible
% states.
% % , based on a set of observations and observation probabilities.
This is a minimal set of assumptions that can be made about an environment faced by an agent engaged in interactive learning.

@@

\paragraph*{Why Markov Decision Processes?} One may wonder whether MDP models
are applicable to something as complex as human behavior. This class of
reinforcement-learning models has had numerous successes in diverse domains. For instance, financial trading is largely a manifestation of strategic decision-making of interacting human agents. According to how the market responds, the agent incurs gain or loss as environmental feedback of the executed financial actions. Recent research on automatizing market exchanges by algorithmic trading has effectively deployed MDPs as a framework for modeling these elaborate behavioral dynamics  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
%
MDPs have also been effective as a behavioral model in robotics
\citep{ng2004,abbeel2004} and in challenging multistep strategy games
~\citep{mnih2015,silver2016mastering,pritzel2017neural}. 
More recent work has
developped an MDP-based way of reasoning about future behavior
of other agents \citep{robinowitz2018}.
The idea is to use meta-learning (learning to learn) to build strong priors
about the behavior of of a population of other agents.

@@

\paragraph*{Reinforcement learning in the brain.}
RL can be plausibly realized in the human brain \citep{doherty2015structure, daw2014algorithmic}.
Indeed,
previous authors have proposed
\citep{gershman2015computational} that a core property of human intelligence is the improvement
of expected utility outcomes as a strategy for action choice in uncertain
environments, a situation perfectly captured by the formalism of MDPs.
It has also long been proposed \citep{dayan2008decision} that there
can be a mapping between algorithmic aspects
underlying model-free and model-based RL learning
and neurobiological aspects underlying brain function.
The neurotransmitter dopamine could serve
as a 'teaching signal' to better estimate value associations
and action policies by modulating
synaptic plasticity in the reward-processing circuitry, including the NAc.
In contrast, model-based RL would start off with some mechanistic assumptions about the dynamics of the world.
These can be assumptions about the physical laws governing the agent's environment or constraints on the state space, transition probabilities between states,
reward contingencies, etc.
% For instance, if a rat in a maze knows
% that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
An agent might represent such knowledge about the world as follows:
\begin{itemize}
\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a location
offering relevant resources.
\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
\item etc.
\end{itemize}
% Model-based assumes knowledge of the environment in terms of the transition probabilities between states, the reward contingencies for state-action pairs, or any form of representation of the environment.
Such knowledge can be partly extracted from the environment: the agent infers a model of the world while learning to take optimal decisions based on the current representation of the environment. These methods learn what the effect is going to be of taking a particular action in a particular state. The result is an estimate of the underlying MDP which can then be either solved exactly or approximately, depending on the setting and what is feasible.

@@

\subsubsection{Accumulated rewards and policies}
The behavior of the agent is governed by a \textit{policy}, which maps states of the world
to probability distributions over candidate actions.
Starting at time $t=0$,
following a policy $\pi$ generates a trajectory of action choices:
\begin{eqnarray*}
  \begin{split}
    \textbf{choose action: }a_0 &\sim \pi(a|s_0)\\
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect
      reward }R_0 = r(s_0, a_0, s_1)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect
      reward }R_1 = r(s_1, a_1,s_2)\\
    \vdots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and
      collect reward }R_{t} = r(s_{t}, a_{t},s_{t+1})\\
    \vdots
  \end{split}
\end{eqnarray*}
We assume time invariance
in that we expect the dynamics of the process
to be equivalent over sufficiently long time windows of equal length (i.e., stationarity).
Since an action executed in the present moment might have repercussions in the
far future, it turns out that the quantity to optimize is not the instantaneous
rewards $r(s, a)$, but a \textit{cumulative reward} estimate which takes into
account expected reward from action choices in the future. A common approach to
modeling this gathered outcome is the time-discounted cumulative reward% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable
  \footnote{Random as it depends both on the environment's dynamics and the
  policy $\pi$ being executed. The exponential delay discounting function used
  here refers to the usual formulation in the field of reinforcement learning,
  although psychological experiments may also reveal other discounting regimes
  \citep{green2004discounting}.}
measures the cumulative reward of
following an action policy $\pi$.
%
Note that value buffering may be realized in the vmPFC by virtue of
this region's connections to the NAc of the reward sytem
\citep{carlezon2009}.

@@

\subsection{The components of reinforcement learning in the DMN}
Given only the limited information available from an MDP, at a state $s$ the average
utility of choosing an action $a$ under a policy $\pi$ can be captured by the single number
% We can now formulate the purpose of the DMN
% as finding a policy $\pi$ for the human agent that maximizes the
% expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a],
\end{equation}
called the $Q$-value for the state-action pair $(s,a)$.
In other words, $Q^{\pi}(s,a)$ corresponds to the expected reward
over all considered action trajectories, in which
the agent sets out in the environment in state
$s$, chooses action $a$, and then follows the policy $\pi$ to select future actions.
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
For the brain,
$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
utility of executing a specific action.
It thus answers the question
``What is the expected utility of choosing action $a$, and its ramifications, in this situation?''.
$Q^{\pi}(s,a)$ offers a formalization of optimal behavior that
may well capture such processing aspects subserved by the DMN in human agents.

@@

Optimal behavior of the agent corresponds to
a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. Formally,
% \textbf{XXX please add 1-2 sentences why this is interesting in the current
% context. This statement appears a little alone as it stands.}
\begin{equation}
  \pi^*(s) := \argmax_{a \in \mathcal A}{Q}^*(s, a),\text{ where }Q^*(s, a) := \max_{\pi}Q^\pi(s,a).
  \label{eq:qlearning}
\end{equation}
The learning goal is to approach the
policy $\pi^*$ as close as possible, that is to \text{solve} the MDP.
Note that \eqref{eq:qlearning} presents merely a definition and
does not lend itself as a candidate schema
for solving MDPs with even moderately sized action and state spaces
(i.e., intractability).
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
Fortunately, the \textit{Bellman equation} \citep{sutton1998reinforcement} provides a fixed-point relation which defines $Q^*$ implicitly via a sampling procedure, without querying the entire space of policies, with the form
\begin{equation}
  Q^* = \text{Bel}(Q^*),
  \label{eq:bellman}
\end{equation}
where the so-called Bellman transform $\text{Bel}(Q)$
of an arbitrary $Q$-value function $Q: \mathcal S \times \mathcal A \rightarrow \mathbb R$  is another $Q$-value function defined by
\begin{equation}
  \begin{split}
   \text{Bel}(Q)(s,a) &:=
   \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q(s', a')]\\
   &= r(s,a) + \gamma\mathbb E_{s' \sim p(s'|s,a)} [\max_{a' \in \mathcal A}Q(s', a')]\\
   &= \text{instantaneous reward} + \text{expected reward for acting greedily thereafter}
%    \; \forall (s,a) \in \mathcal S \times \mathcal A
    % \\&=
    % \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^*(s', \pi^*(s'))].
    \end{split}
  \end{equation}

@@

% \subsection{Learning in the DMN}
\subsubsection{Value approximation and the policy matrix}
\label{sec:policymat}
As already mentioned in the previous section, Q-learning~\citep{watkins92} optimizes over the class of
deterministic policies of the form \eqref{eq:qlearning}. State spaces may be extremely large
and tracking all possible states and actions may require prohibitively excessive
computation and memory resources.
The need of maintaining an explicit table of
states can be eliminated by instead using of an approximate $Q$-value function $\tilde{Q}(s,a|\theta)$
by keeping track of an approximating parameter $\theta$ of much lower dimension than the number of states.
At a given time step, the world is in a state $s \in \mathcal S$, and the agent takes an
action which it expects to be the most valuable on average, namely

\begin{equation}
  \pi^{\text{hard-max}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a|\theta).
  \label{eq:policy}
\end{equation}
This defines a mapping from states directly to actions.
% $a \mapsto \tilde{Q}(a,s|\theta)$ of the $Q$-value function ``frozen'' at state $s$.
For instance, a simple linear model with a kernel $\phi$ would be of the
form $\tilde{Q}(s, a|\theta) = \phi(s,a)^T\theta$, where
$\phi(s,a)$ would represent a high-level representation of the state-action pairs
$(s,a)$, as was previously proposed \citep{songNIPS2016},
or artificial neural-network models as demonstrated in recent
seminal investigations
\citep{mnih2015,silver2016mastering} for playing complex games (atari, Go, etc.) at super-human levels.
%  and Parr et al. 2003 (JMLR)].
In the DMN, the dmPFC would implement such a hard-max lookup
over the action space.
The model
parameters $\theta$ would correspond to synaptic weights and connection strengths within and between
brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the hard-max policy \eqref{eq:policy}.
The approximating $Q$-value function $\tilde{Q}(s, a|\theta)$ would inform the DMN with the (expected) usefulness of choosing an action $a$ in state $s$.
The DMN, and in particular its dmPFC part, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes the approximate
$Q$-values.
This mapping from states to actions that is conventionally called
\textit{policy matrix} \citep{mnih2015,silver2016mastering}.
Learning consists in starting from a given table and
updating it during action choices,
which take the agent to different table entries.

@@

\subsubsection{Does the hippocampus subserve Monte Carlo sampling?}
In RL, Monte Carlo simulation is a common means to update the agent's belief state
based on stochastic sampling around states and possible transitions \citep{daw2014algorithmic,silver2010monte}.
Monte Carlo simulation provides a simple method for evaluating the value of a state.
This inference procedure provides an effective mechanism both for tree search (of the considered
action trajectories)
and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than a RL agent without stochastic resampling procedures.
Such methods have scaling as a function of available data (i.e., sample complexity) that
is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
which can be prohibitively large.

@@

\paragraph*{A note on implicit and explicit memory.}
While Markov processes are usually memoryless, it is mathematically feasible to
incorporate the previous states of such model into the current state. This
extension may partially account for implicit memory at the behavioral level, but
may not explain the underlying neurobiological implementation or accommodate
explicit memory. Implicit memory-based processing arises in our MDP account of
DMN function in several different forms: successive updates of a) the action
policy and the value function, both being products of the past, as well as b)
the deep non-linear relationships within the hierarchical connections of
biological neural networks. The brain's adaptive synaptic connections can be
thought as a deep neural-network architecture affording an implicit form of
information compression of life experience. Such memory traces are stored in the
neural machinery and can be implicitly retrieved as a form of knowledge during
simulation of action rather than accessed as a stored explicit representation
\citep{pezzulo2011grounding}. c) Certain neural processes in the hippocampus can
be seen as some type of Monte Carlo sampling for memory recall, which can also
be a basis for probabilistic simulations across time scales
\citep{schacter2007remembering,Axelrod2017}.

\subsection{Putting everything together}

The DMN is today known to consistently increase in neural
activity when humans engage in cognitive processes that are relatively detached from
the current sensory environment.
The more familiar and predictable the current environment, the more brain resources may remain for allocating DMN activity to MDP processes extending beyond the present time and sensory context.
%
This speculation receives quantitative support in that connectional links between
nodes of the DMN have been reported to be more consistent and reliable than
functional couplings within any other macroscopical networks \citep{shehz2009}.
As such, random-sampling-related baseline evaluation of possibilities and
events may be subserved by the DMN and get partly suspended when novelty in
the external environment is encountered or immediate action is required.
%
In line with this perspective, DMN engagement was shown to heighten and relate to effective behavioral responses in the practiced phase of a demanding cognitive flexibility task, as compared to acquisition phase when participants learned context-specific rules. This involvement in automated decision-making has led the authors to propose an ``autopilot'' role for the DMN \citep{vatansever2017pnas}, which may contribute to optimizing intervention of the organism on the world in general. Among all parts of the DMN, the RTPJ is perhaps the most evident candidate for a network-switching relay that calibrates between processing of environment-engaged versus internally generated information \citep{downar2000multimodal, golland2006extrinsic, bzdok2013tpj}.

>>>>

RL is an area of machine learning concerned with searching optimal behavioral strategies through interactions
with an \textit{environment} with the goal to maximize the \textit{cumulative reward} over time.
The optimal behavior typically
takes the future into account as certain rewards could be \textit{delayed}.
Through repeated action on and feedback from the environment,
the agent learns how to reach goals and continuously improve the collection of reward signals
in a trial-and-error fashion (Fig. \ref{fig:rl}).
At a given moment, each taken \textit{action} $a$ triggers a change
in the \textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
$r = r(s, a,s')$ obtained by the agent.
If the collected reward outcome yields a negative value it can be
more naturally interpreted as \textit{punishment}.
The environment be partly controlled by
the action of the agent and the reward can be thought
of as satisfaction \textemdash or aversion \textemdash that accompany the execution of
a particular action.
% Action may be delayed to achieve substantial increases in expected reward
% that is expected to grow with time.

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory input
\citep{starkweather2017dopamine}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that an abstract description of DMN activity based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.
In short, DMN implication in a diversity of
particularly sophisticated human behaviors
can be parsimoniously explained as instantiating probabilistic
simulations of experience coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time scales.
Such a purposeful optimization objective
may be subserved by a stochastic approximation
based on a brain implementation for Monte Carlo sampling of events and outcomes.

@@

\subsection{Markov decision processes}
% RL has had considerable success in modeling complex decision-making problems,
% We emphasize that the brain in general, and the DMN in particular,
% is a physical system governed by the laws of
% physics and can be formally described
% by Markov processes at a sufficiently coarse scale.
% It has indeed been previously proposed \citep{tegmark2016improved} that
% any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
% step is sufficiently short.
In artificial intelligence and machine learning, a popular computational model for
multi-step decision processes are
MDPs \citep{sutton1998reinforcement}.
% For the remainder of this paper, for the sake of simplicity
% we use the abbreviation ``MDP'' to refer to the special case
% of \textit{Partially Observable Markov Decision Processes}.
An MDP operationalizes a sequential decision process
in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state.
Instead, the agent tries to optimize a \textit{subjective} reward
signal (i.e., likely to be different for another agent in the same state
and possibly driven by neural processing in the vmPFC)
by maintaining probability distributions over actions according to their expected utility.
% This is performed alongside each transition in the state space of the environment
%  and maintains a probability distribution over the set of possible
% states.
% % , based on a set of observations and observation probabilities.
This is a minimal set of assumptions that can be made about an environment faced by an agent engaged in interactive learning.

@@

\paragraph*{Why Markov Decision Processes?} One may wonder whether MDP models
are applicable to something as complex as human behavior. This class of
reinforcement-learning models has had numerous successes in diverse domains. For instance, financial trading is largely a manifestation of strategic decision-making of interacting human agents. According to how the market responds, the agent incurs gain or loss as environmental feedback of the executed financial actions. Recent research on automatizing market exchanges by algorithmic trading has effectively deployed MDPs as a framework for modeling these elaborate behavioral dynamics  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
%
MDPs have also been effective as a behavioral model in robotics
\citep{ng2004,abbeel2004} and in challenging multistep strategy games
~\citep{mnih2015,silver2016mastering,pritzel2017neural}. 
More recent work has
developped an MDP-based way of reasoning about future behavior
of other agents \citep{robinowitz2018}.
The idea is to use meta-learning (i.e., learning to learn) to build strong priors
about the behavior of a population of other agents.

@@

\paragraph*{Reinforcement learning in the brain.}
RL has been argued to be a biologically plausible mechanism in the human brain
\citep{doherty2015structure, daw2014algorithmic}.
Indeed,
previous authors have proposed
\citep{gershman2015computational} that a core property of human intelligence is the improvement
of expected utility outcomes as a strategy for action choice in uncertain
environments, a situation captured by the formalism of MDPs.
It has also long been proposed \citep{dayan2008decision} that there
can be a mapping between algorithmic aspects
underlying model-free and model-based RL learning
and neurobiological aspects underlying brain function.
The neurotransmitter dopamine could serve
as a 'teaching signal' to guide estimation of value associations
and action policies by modulating
synaptic plasticity in the reward-processing circuitry, including the NAc.
In contrast, model-based RL would start off with some mechanistic assumptions about the dynamics of the world.
These assumptions could relate to the
physical laws governing the agent's environment, constraints on the state space, transition probabilities between states, or reward contingencies.
% For instance, if a rat in a maze knows
% that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
An agent might represent such knowledge about the world as follows:
\begin{itemize}
\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a location
offering relevant resources.
\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
\item etc.
\end{itemize}
% Model-based assumes knowledge of the environment in terms of the transition probabilities between states, the reward contingencies for state-action pairs, or any form of representation of the environment.
Such knowledge can be partly extracted from the environment: the agent infers a model of the world while learning to take optimal decisions based on the current representation of the environment. These methods learn what the effect is going to be of taking a particular action in a particular state. The result is an estimate of the underlying MDP which can then be either solved exactly or approximately, depending on the setting and what is feasible.

@@

\subsubsection{Accumulated rewards and policies}
The behavior of the agent is governed by a \textit{policy}, which maps states of the world
to probability distributions over candidate actions.
Starting at time $t=0$,
following a policy $\pi$ generates a trajectory of action choices:
\begin{eqnarray*}
  \begin{split}
    \textbf{choose action: }a_0 &\sim \pi(a|s_0)\\
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect
      reward }R_0 = r(s_0, a_0, s_1)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect
      reward }R_1 = r(s_1, a_1,s_2)\\
    \vdots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and
      collect reward }R_{t} = r(s_{t}, a_{t},s_{t+1})\\
    \vdots
  \end{split}
\end{eqnarray*}
We assume time invariance
in that we expect the dynamics of the process
to be equivalent over sufficiently long time windows of equal length (i.e., stationarity).
Since an action executed in the present moment might have repercussions in the
far future, it turns out that the quantity to optimize is not the instantaneous
rewards $r(s, a)$, but a \textit{cumulative reward} estimate which takes into
account expected reward from action choices in the future. A common approach to
modeling this gathered outcome is the time-discounted cumulative reward% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable
  \footnote{The reward outcome is random as it depends both on the environment's dynamics and the
  policy $\pi$ being executed. The exponential delay discounting function used
  here refers to the usual formulation in the field of reinforcement learning,
  although psychological experiments may also reveal other discounting regimes
  \citep{green2004discounting}.}
measures the cumulative reward of
following an action policy $\pi$.
%
Note that value buffering may be realized in the vmPFC by virtue of
this region's connections to the NAc of the reward sytem
\citep{carlezon2009}.

@@

\subsection{The components of reinforcement learning in the DMN}
Given only the limited information available from an MDP, at a state $s$ the average
utility of choosing an action $a$ under a policy $\pi$ can be captured by the single number
% We can now formulate the purpose of the DMN
% as finding a policy $\pi$ for the human agent that maximizes the
% expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a],
\end{equation}
called the $Q$-value for the state-action pair $(s,a)$.
In other words, $Q^{\pi}(s,a)$ corresponds to the expected reward
over all considered action trajectories, in which
the agent sets out in the environment in state
$s$, chooses action $a$, and then follows the policy $\pi$ to select future actions.
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
For the brain,
$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
utility of executing a specific action.
In this way, we can answer the question
``What is the expected utility of choosing action $a$, and its ramifications, in this situation?''.
$Q^{\pi}(s,a)$ offers a formalization of optimal behavior that
may well capture processing aspects such as subserved by the DMN in human agents.

@@

Optimal behavior of the agent corresponds to
a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. Formally,
% \textbf{XXX please add 1-2 sentences why this is interesting in the current
% context. This statement appears a little alone as it stands.}
\begin{equation}
  \pi^*(s) := \argmax_{a \in \mathcal A}{Q}^*(s, a),\text{ where }Q^*(s, a) := \max_{\pi}Q^\pi(s,a).
  \label{eq:qlearning}
\end{equation}
The learning goal is to approach the
policy $\pi^*$ as close as possible, that is to \text{solve} the MDP.
Note that \eqref{eq:qlearning} presents merely a definition and
does not lend itself as a candidate schema
for solving MDPs with even moderately sized action and state spaces
(i.e., computational intractability).
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
Fortunately, the \textit{Bellman equation} \citep{sutton1998reinforcement} provides a fixed-point relation which defines $Q^*$ implicitly via a sampling procedure, without querying the entire space of policies, with the form
\begin{equation}
  Q^* = \text{Bel}(Q^*),
  \label{eq:bellman}
\end{equation}
where the so-called Bellman transform $\text{Bel}(Q)$
of an arbitrary $Q$-value function $Q: \mathcal S \times \mathcal A \rightarrow \mathbb R$  is another $Q$-value function defined by
\begin{equation}
  \begin{split}
   \text{Bel}(Q)(s,a) &:=
   \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q(s', a')]\\
   &= r(s,a) + \gamma\mathbb E_{s' \sim p(s'|s,a)} [\max_{a' \in \mathcal A}Q(s', a')]\\
   &= \text{instantaneous reward} + \text{expected reward for acting greedily thereafter}
%    \; \forall (s,a) \in \mathcal S \times \mathcal A
    % \\&=
    % \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^*(s', \pi^*(s'))].
    \end{split}
  \end{equation}

@@

% \subsection{Learning in the DMN}
\subsubsection{Value approximation and the policy matrix}
\label{sec:policymat}
As already mentioned in the previous section, Q-learning~\citep{watkins92} optimizes over the class of
deterministic policies of the form \eqref{eq:qlearning}. State spaces may be extremely large
and tracking all possible states and actions may require prohibitively excessive
computation and memory resources, perhaps reflect in the especially high metabolic
turn-over of the posterior medial DMN (PMC).
The need of maintaining an explicit table of
states can be eliminated by instead using of an approximate $Q$-value function $\tilde{Q}(s,a|\theta)$
by keeping track of an approximating parameter $\theta$ of much lower dimension than the number of states.
At a given time step, the world is in a state $s \in \mathcal S$, and the agent takes an
action which it expects to be the most valuable on average, namely

\begin{equation}
  \pi^{\text{hard-max}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a|\theta).
  \label{eq:policy}
\end{equation}
This defines a mapping from states directly to actions.
% $a \mapsto \tilde{Q}(a,s|\theta)$ of the $Q$-value function ``frozen'' at state $s$.
For instance, a simple linear model with a kernel $\phi$ would be of the
form $\tilde{Q}(s, a|\theta) = \phi(s,a)^T\theta$, where
$\phi(s,a)$ would represent a high-level representation of the state-action pairs
$(s,a)$, as was previously proposed \citep{songNIPS2016},
or artificial neural-network models as demonstrated in 
seminal machine-learning models
\citep{mnih2015,silver2016mastering} for playing complex games (atari, Go, etc.) at super-human levels.
%  and Parr et al. 2003 (JMLR)].
In the DMN, the dmPFC would implement such a hard-max lookup
over the action space.
The model
parameters $\theta$ would correspond to synaptic weights and connection strengths within and between
brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the hard-max policy \eqref{eq:policy}.
The approximating $Q$-value function $\tilde{Q}(s, a|\theta)$ would inform the DMN with the (expected) usefulness of choosing an action $a$ in state $s$.
The DMN, and in particular its dmPFC part, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes the approximate
$Q$-values.
This mapping from states to actions that is conventionally called
\textit{policy matrix} \citep{mnih2015,silver2016mastering}.
Learning consists in starting from a given table and
updating it during action choices, potentially reflected in neural processing in the PMC,
which take the agent to different table entries.

@@

\subsubsection{Does the hippocampus subserve Monte Carlo sampling?}
In RL, Monte Carlo simulation is a common means to update the agent's belief state
based on stochastic sampling around states and possible transitions \citep{daw2014algorithmic,silver2010monte}.
Monte Carlo simulation provides a simple method for evaluating the value of a state.
This inference procedure provides an effective mechanism both for tree search (of the considered
action trajectories)
and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than a RL agent without stochastic resampling procedures.
Such methods scale as a function of available data (i.e., sample complexity) that
is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
which can be prohibitively large.

@@

\paragraph*{A note on implicit and explicit memory.}
While Markov processes are usually memoryless, it is mathematically feasible to
incorporate a set of previous states of such model into the current state. This
extension may partly account for implicit memory at the behavioral level, but
may not explain the underlying neurobiological implementation or accommodate
explicit memory. Implicit memory-based processing arises in our MDP account of
DMN function in several different forms: successive updates of a) the action
policy and the value function, both being products of the past, as well as b)
the deep non-linear relationships within the hierarchical connections of
biological neural networks (especially in the association cortex).
The brain's adaptive synaptic connections can be
viewed as a deep neural-network architecture affording an implicit form of
information compression of life experience. Such memory traces are stored in the
neural machinery and can be implicitly retrieved as a form of knowledge during
simulation of action rather than accessed as a stored explicit representation
\citep{pezzulo2011grounding}. c) Certain neural processes in the hippocampus can
be seen as some type of Monte Carlo sampling for memory recall, which can also
be a basis for probabilistic simulations across time scales
\citep{schacter2007remembering,Axelrod2017}.

\subsection{Putting everything together: hypotheses for future studies}

The DMN is today known to consistently increase in neural
activity when humans engage in cognitive processes that are relatively detached from
the current sensory environment.
The more familiar and predictable the current environment, the more brain resources may remain for allocating DMN activity to MDP processes extending beyond the present time and sensory context.
%
This speculation receives quantitative support in that connectional links between
nodes of the DMN have been reported to be more consistent and reliable than
functional couplings within any other macroscopical networks \citep{shehz2009}.
As such, random-sampling-related baseline evaluation of action possibilities and
their consequences may be subserved by the DMN and get partly suspended when novelty in
the external environment is encountered or immediate action is required.
%
In line with this perspective, DMN engagement was shown to heighten and relate to effective behavioral responses in the practiced phase of a demanding cognitive flexibility task, as compared to acquisition phase when participants learned context-specific rules. This involvement in automated decision-making has led the authors to propose an ``autopilot'' role for the DMN \citep{vatansever2017pnas}, which may contribute to optimizing intervention of the organism on the world in general. Among all parts of the DMN, the RTPJ is perhaps the most evident candidate for a network-switching relay that calibrates between processing of environment-engaged versus internally generated information \citep{downar2000multimodal, golland2006extrinsic, bzdok2013tpj}.
